### 2016-05-09 bioc192/lect04.tex
tmp commit

In concerted model, all of the subunits are either in the T state, or the R state, and the states are at equilibrium (i.e. no `in between' state).
The O$_2$ molecules can bind to either state at any time.
When there are more substrate (i.e. O$_2$) is bound to haemoglobin, the greater the chance for the subunits to be in the R state, which means that the equilibrium is favoured towards the R state when there are more O$_2$ molecules bound to haemoglobin.

\textbf{ In concerted model, the T and R states are at equilibrium and substrates can bind at any time to any conformational state. The binding of substrate causes the equilibrium to shift to the R state and the substrates are more likely to bind to the protein.}

>>>>

In concerted model, all of the subunits are either in the T state, or the R state, and the states are at equilibrium (i.e.\  no `in between' state).
The O$_2$ molecules can bind to either state at any time.
When there are more substrate (i.e. O$_2$) is bound to haemoglobin, the greater the chance for the subunits to be in the R state, which means that the equilibrium is favoured towards the R state when there are more O$_2$ molecules bound to haemoglobin.

\textbf{In concerted model, the T and R states are at equilibrium and substrates can bind at any time to any conformational state. The binding of substrate causes the equilibrium to shift to the R state and the substrates are more likely to bind to the protein.}

### 2016-05-09 bioc192/lect05.tex
tmp commit

2,3-Bisphosphoglycerate (BPG) is a small, negatively charged (-5 charge) molecule that binds to the deoxy-haemoglobin at the center of the tetramer.
Binding of BPG causes the haemoglobin to be in T state, i.e. in the dexygenated state, and decreases the haemoglobin's binding affinity for O$_2$, meaning that the haemoglobin with BPG bound to it will require higher pO$_2$ to oxygenate it.
Binding of BPG at the muscle/tissue site ensures that the O$_2$ molecules are released at the muscles/tissues.

@@

Both H$^+$ and CO$_2$ binds to haemoglobin at a different site as the substrate (O$_2$) and affects the oxygen affinity of haemoglobin.
Both molecules favour the T state of haemoglobin, and therefore favours the deoxygenated state (i.e. decreased O$_2$ affinity).
Note that both of these molecules are related to the pH of the environment, and the blood buffer system is also involved in controlling the concentration of these molecules.

@@

Technically speaking, the Bohr effect is the effect of pH (i.e. [H$^+$]) and [CO$_2$] on the oxygen affinity of haemoglobin.
As pH decreases (therefore [H$^+$] increases), the net charge of haemoglobin changes due to different amino acids getting protonated, which leads to slight conformational changes in the haemoglobin structure, and therefore decrease in haemoglobin's oxygen affinity.
Similarly, as CO$_2$ concentration increases, CO$_2$ is more likely to bind to the haemoglobin and decrease the oxygen affinity of haemoglobin.

@@

Also note that BPG is released at the lungs when O$_2$ binds to the haemoglobin, and binds at the tissue/muscle when O$_2$ is released and CO$_2$ and H$^+$ binds to it, contributing to the binding and release of O$_2$.(insert the saturation curve of +/- allosteric effectors)

@@

\begin{itemize}
	\item Allosteric effectors are molecules that bind to the allosteric protein at a place other than/different to the normal substrate binds to.
	\item
	\item
	\item
	\item
	\item
	\item
	\item
	\item
\end{itemize}

>>>>

2,3-Bisphosphoglycerate (BPG) is a small, negatively charged (-5 charge) molecule that binds to the deoxy-haemoglobin at the center of the tetramer.
Binding of BPG causes the haemoglobin to be in T state, i.e.\  in the dexygenated state, and decreases the haemoglobin's binding affinity for O$_2$, meaning that the haemoglobin with BPG bound to it will require higher pO$_2$ to oxygenate it.
Binding of BPG at the muscle/tissue site ensures that the O$_2$ molecules are released at the muscles/tissues.

@@

Both H$^+$ and CO$_2$ binds to haemoglobin at a different site as the substrate (O$_2$) and affects the oxygen affinity of haemoglobin.
Both molecules favour the T state of haemoglobin, and therefore favours the deoxygenated state (i.e.\  decreased O$_2$ affinity).
Note that both of these molecules are related to the pH of the environment, and the blood buffer system is also involved in controlling the concentration of these molecules.

@@

Technically speaking, the Bohr effect is the effect of pH (i.e.\  [H$^+$]) and [CO$_2$] on the oxygen affinity of haemoglobin.
As pH decreases (therefore [H$^+$] increases), the net charge of haemoglobin changes due to different amino acids getting protonated, which leads to slight conformational changes in the haemoglobin structure, and therefore decrease in haemoglobin's oxygen affinity.
Similarly, as CO$_2$ concentration increases, CO$_2$ is more likely to bind to the haemoglobin and decrease the oxygen affinity of haemoglobin.

@@

Also note that BPG is released at the lungs when O$_2$ binds to the haemoglobin, and binds at the tissue/muscle when O$_2$ is released and CO$_2$ and H$^+$ binds to it, contributing to the binding and release of O$_2$ (insert the saturation curve of +/- allosteric effectors).

@@

\begin{itemize}
	\item Allosteric effectors are molecules that bind to the allosteric protein at a place other than/different to where the normal substrate binds to
	\item BPG is a negatively charged small molecule that binds at the center of the haemoglobin tetramer and favours the T state
	\item Both H$^+$ and CO$_2$ are allosteric effectors that favours the T state of haemoglobin
	\item H$^+$ and CO$_2$ are closely related to the pH of the environment, so these molecules will have different effect in different parts of the body
\end{itemize}

### 2016-05-09 bioc192/lect06.tex
added graphics
editing bioc notes
tmp commit

@@

Haemoglobin C (HbC) is common in West Africa, and is caused by a point mutation that causes a substitution mutation.
The substitution is at the same place as HbS: glutamic acid to lysine at position 6 (E6K).
As a consequence of this substitution, the charge is swapped from negatively charged glutamic acid to positively charged lysine.

@@

>>>>

\begin{center}
% \includegraphics[width=0.5\textwidth]{hbs}
\end{center}

@@

Haemoglobin C (HbC) is common in West Africa, and is caused by a point mutation that causes a substitution mutation.
The substitution is at the same place as HbS\@ glutamic acid to lysine at position 6 (E6K).
As a consequence of this substitution, the charge is swapped from negatively charged glutamic acid to positively charged lysine.

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Remember the haemoglobin variants mentioned in this note in some details
\end{itemize}

### 2016-05-09 bioc192/lect07.tex
added graphics
editing bioc notes
tmp commit

Most proteins, if not all proteins, in eukaryotic cells are post-translationally modified.
Post-translational modification increases the number of distinct forms of proteins without increasing the number of genes in the genome.
This means that the level of complexity provided by the proteins are also increased as well.

The diversity of the proteins can be expanded by mRNA splicing and PTM.
PTM can be categorised into two categories: the covalent addition of chemical group to amino acid side chain, or covalent hydrolysis of the peptide bond in a protein.

@@

\section{Phoshphorylation}

@@

Quite often, phosphorylation state is used to `switch' the conformational state of the protein to enable a biological process in a cell (e.g. facilitating interaction with other proteins).

\subsection{Insulin Receptor}

Binding of insulin to the insulin receptor causes a conformational change in the receptor, and this exposes the tyrosine in the cytosol side of the membrane.
The exposed tyrosine is then targeted by kinases, and get phosphorylated by it.
Phosphorylated tyrosine can then cause responses (e.g. signal transduction, GLUT4 transporter expression).

\subsection{Na$^+$/K$^+$ Pump}

@@

\section{Gamma-carboxy Glutamic Acid (Gla)}

@@

\section{Hydroxylation}

@@

>>>>

Most, if not all,  proteins in eukaryotic cells are post-translationally modified.
Post-translational modification increases the number of distinct forms of proteins without increasing the number of genes in the genome.
This means that the level of complexity provided by the proteins are also increased as well.

The diversity of the proteins can be expanded by mRNA splicing and PTM\@ (mRNA splicing will not be covered).
PTM can be categorised into two categories: the covalent addition of chemical group to amino acid side chain, or covalent hydrolysis of the peptide bond in a protein.

@@

\subsection{Phoshphorylation}

@@

Quite often, phosphorylation state is used to `switch' the conformational state of the protein to enable a biological process in a cell (e.g.\ facilitating interaction with other proteins).

\subsubsection{Insulin Receptor}

Binding of insulin to the insulin receptor causes a conformational change in the receptor, and this exposes the tyrosine in the cytosol side of the membrane.
The exposed tyrosine is then targeted by kinases, and get phosphorylated by it.
Phosphorylated tyrosine can then cause responses (e.g.\ signal transduction, GLUT4 transporter expression).

\begin{center}
% \includegraphics[width=0.7\textwidth]{insulinreceptor}
\end{center}

\subsubsection{Na$^+$/K$^+$ Pump}

@@

\begin{center}
% \includegraphics[width=0.7\textwidth]{nakpump}
\end{center}

\subsection{Gamma-carboxy Glutamic Acid (Gla)}

@@

\subsection{Hydroxylation}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Phosphorylation occurs on the hydroxyl group of the amino acid side chains
	\item Kinases phosphorylates amino acids, and phosphatases dephosphorylates amino acids
	\item Phosphorylation results in an addition of a large negatively charged molecule to the protein
	\item Gamma-carboxy glutamic acid has an extra corboxyl group added onto glutamic acid
	\item Gamma-carboxy Glutamic Acid is done in some blood co-agulation proteins
	\item Hydroxylation is an addition of the hydroxyl group to proline or lysine
	\item Hydroxylation provides greater possibilities of hydrogen bonding with other atoms, and therefore provides strength/order in structures like collagen
\end{itemize}

### 2016-05-09 bioc192/lect09.tex
added graphics
editing bioc notes
tmp commit

Therapeutic proteins were extracted from relevant sources (e.g. tissues, serum) and was used to treat the patients.
The problems with this approach was the risk of identified and unidentified pathogen infections from the source (e.g. HIV).

@@

Recombinant antibodies are designed to control the effects of some proteins in your body (e.g. some cell receptors).
Mouse AB can be made, but it is recognised as foreign material in humans and get rejected.
This lead to the formation of chimeric antibodies, so there will be no immune response.

@@

>>>>

Therapeutic proteins were extracted from relevant sources (e.g.\ tissues, serum) and was used to treat the patients.
The problems with this approach was the risk of identified and unidentified pathogen infections from the source (e.g. HIV).

@@

Recombinant antibodies are designed to control the effects of some proteins in your body (e.g.\ some cell receptors).
Mouse AB can be made, but it is recognised as foreign material in humans and get rejected.
This lead to the formation of chimeric antibodies, so there will be no immune response.

\begin{center}
% \includegraphics[width=0.8\textwidth]{chimericab}
\end{center}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Read through this lecture section notes
\end{itemize}

### 2016-05-09 bioc192/lect10.tex
added topic2 graphics
editing bioc notes
tmp commit

Geometric specificity is where the substrate must be made of same component/structure for it to be recognised by the enzyme.
Stereospecificity is where the substrate has the required components,but don't have the right orientation, and therefore not recognised by the enzymes.

@@

\begin{center}
\ch{ E + S <=> ES <=> [{T}] EP <=> E + P }
\end{center}

@@

@@

@@

>>>>

\begin{center}
% \includegraphics[width=0.7\textwidth]{geo_stereo}
\end{center}
Geometric specificity is where the substrate must be made of same component/structure for it to be recognised by the enzyme.
Stereospecificity is where the substrate has the required components,but don't have the right orientation, and therefore not recognised by the enzymes.

@@

\begin{center}
\ch{E + S <=> ES <=> [{T}] EP <=> E + P}
\end{center}

@@

\begin{center}
% \includegraphics[width=0.6\textwidth]{enzymemodels}
\end{center}

@@

\begin{center}
% \includegraphics[width=0.5\textwidth]{enzcatalreac}
\end{center}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Geometric specificity refers to the enzyme being specific  to the component/structure of the substrate
	\item Stereospecificity refers to the enzyme being specific to the orientation of the substrate
	\item Enzyme is not altered after an enzymatic reaction
	\item Lock-and-key model is where the substrate perfectly matches the enzyme's active site
	\item Induced-fit model is where the binding of the substrate to the enzyme causes a conformational change in both the substrate and the enzyme
	\item Know the six catalytic mechanisms of an enzymatic reaction
	\item Co-factors are molecules that are used in the enzymatic reaction, but is changed after the reaction completes
	\item Know the ``numbers'' associated with each of the enzyme classes
\end{itemize}

### 2016-05-09 bioc192/lect11.tex
added topic2 graphics
editing bioc notes
tmp commit

@@

Consider the situation where there is an excess substrate (i.e. unlimited source of substrate).
As the concentration of the enzyme increases, the rate of reaction increases.
\textbf{This means that the velocity of the reaction (V) is proportional to the enzyme concentration, when there is excess substrate.}

@@

Consider the situation where the concentration of enzyme is fixed.
As we increase the concentration of substrate, the rate of reaction will initially increase linearly, but as more and more substrates are introduced, the rate of reaction reaches a `maximum'.
This is because all of the active site of the enzymes become occupied as the substrate concentration increases -- it reaches the `maximum capacity' of the enzyme, and the enzymes cannot increase the rate of reaction no matter how hard it tries.

@@

>>>>

\begin{center}
% \includegraphics[width=0.5\textwidth]{progresscurve}
\end{center}

@@

\begin{center}
% \includegraphics[width=0.5\textwidth]{vvse}
\end{center}

Consider the situation where there is an excess substrate (i.e.\ unlimited source of substrate).
As the concentration of the enzyme increases, the rate of reaction increases.
\textbf{This means that the velocity of the reaction (V) is proportional to the enzyme concentration, when there is excess substrate.}

@@

\begin{center}
% \includegraphics[width=0.4\textwidth]{vvss}
\end{center}

Consider the situation where the concentration of the enzyme is fixed.
As we increase the concentration of substrate, the rate of reaction will initially increase linearly, but as more and more substrates are introduced, the rate of reaction reaches a `maximum'.
This is because all of the active site of the enzymes become occupied as the substrate concentration increases -- it reaches the `maximum capacity' of the enzyme, and the enzymes cannot increase the rate of reaction no matter how hard it tries.

\begin{center}
% \includegraphics[width=0.4\textwidth]{zerofirstkinetics}
\end{center}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Progress curve shows how much product is made over time
	\item Only the initial part of the curve is used, so other factors don't limit the reaction as it proceeds
	\item Reaction velocity is proportional to the enzyme concentration when there is excess substrate
	\item First-order kinetics is where the rate of reaction is proportional to the concentration of the substrate
	\item zero-order kinetics is where the rate of reaction is constant with changing substrate concentration
\end{itemize}

### 2016-05-09 bioc192/lect12.tex
finished lecture 13, onto lecture 14
added topic2 graphics
editing bioc notes
tmp commit

@@

\section{V$_{max}$ k$_{cat}$}

@@

>>>>

\begin{center}
% \includegraphics[width=0.8\textwidth]{lineweaverburk}
\end{center}

@@

\section{V$_{max}$ and k$_{cat}$}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item K$_M$ is defined as the substrate concentration at which the velocity of the reaction is half the maximum velocity
	\item K$_M$ is also defined as the ratio of the breakdown and the formation of the enzyme-substrate complex
	\item K$_M$ characterises a single enzyme-substrate pair, meaning that the value of K$_M$ is different for every substrate and/or enzyme
	\item V$_{max}$ is defined as the maximum velocity of the reaction when an enzyme is saturated with the substrate
	\item Know the Michaelis-Menten equation, and its assumptions
	\item Lineweaver-Burk plot is the V vs [S] curve
	\item k$_{cat}$ is the number of moles of substrate converted into product per mole of enzyme per second
	\item Catalytic efficiency of an enzyme is measured as the ratio of k$_{cat}$ and K$_M$ (i.e.\ how quickly a product is formed relative to how well the substrate binds to the enzyme)
\end{itemize}

### 2016-05-09 bioc192/lect13.tex
added topic2 graphics
editing bioc notes
tmp commit

@@

@@

@@

>>>>

\begin{center}
% \includegraphics[width=0.6\textwidth]{compinh}
\end{center}

@@

\begin{center}
% \includegraphics[width=0.6\textwidth]{noncompinh}
\end{center}

@@

\begin{center}
% \includegraphics[width=0.6\textwidth]{allosterism}
\end{center}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Irreversible inhibitors covalently binds to the enzyme and permanently inactivates it
	\item Competitive reversible inhibitors increases the K$_M$ but doesn't affect the V$_{max}$ of the enzyme
	\item Non-competitive reversible inhibitors decreases the V$_{max}$ but doesn't affect the K$_M$ of the enzyme
\end{itemize}

### 2016-05-09 bioc192/lect14.tex
added images for topic 3
editing bioc notes
tmp commit

@@

@@

@@

@@

>>>>

\begin{center}
% \includegraphics[width=0.4\textwidth]{physrec1}
% \includegraphics[width=0.4\textwidth]{physrec2}
\end{center}

@@

\begin{center}
% \includegraphics[width=1.0\textwidth]{fourreceptors}
\end{center}

@@

\begin{center}
% \includegraphics[width=0.7\textwidth]{ligandgated}
\end{center}

@@

\begin{center}
% \includegraphics[width=1.0\textwidth]{voltagegated}
\end{center}

@@

\subsection*{What you should get out of this lecture}

\begin{itemize}
	\item Drug is a chemical that changes the behaviour of the cell
	\item Agonists are the drugs or molecules that causes the full biological response
	\item Antagonists are the drugs or molecules that prevents the full biological response (opposite of an agonist)
	\item Drug receptor is anything that a drug binds to (DNA/RNA/protein/peptide)
	\item Physiological receptor refers specifically to the  receptors located on the cell surface
	\item Know how the gated channels work
\end{itemize}
